\bf Adaption of the Chumbley Score to matching of bullet striation marks

### 2017-12-26 doc/paper.tex
modification to algorithm in the same shift step..search vicinity

---
title: "Adaption of the Chumbley Score to matching of bullet striation marks"
authors:
- affiliation: Department of Statistics, Iowa State University 
  name: Ganesh Krishnan
  thanks: The authors gratefully acknowledge ...
- affiliation: Department of Statistics and csafe, Iowa State University 
  name: Heike Hofmann
biblio-style: apsr
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    keep_tex: yes
    latex_engine: pdflatex
    template: template.tex
  html_document: default
blinded: 0
keywords:
- 3 to 6 keywords
- that do not appear in the title
bibliography: bibliography
abstract: null
---

@@

\tableofcontents

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
library(knitr)
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  error = FALSE, 
  collapse = TRUE,
  comment = "",
  fig.height = 6,
  fig.width = 6,
  fig.align = "center",
  out.width= '\\textwidth',
  cache = FALSE,
  fig.path='figures/',
  echo=FALSE,
  cache=TRUE
)
options(knitr.table.format = "latex")
library(tidyverse)
```

```{r functions}
errorrate <- function(data, alpha) {
  summ <- data %>% filter(!is.na(p_value)) %>%
    mutate(signif = p_value < alpha) %>%
    group_by(wv, wo, match, coarse, signif) %>% tally()
  summ$error <- with(summ, match != signif)
  summ$alpha <- alpha

  rates <- summ %>% group_by(wv, wo, coarse, match, alpha) %>% summarize(
    rate = n[error==TRUE]/sum(n)
  )
  totals <- summ %>% group_by(wv, wo, coarse, alpha) %>% summarize(
    total= sum(n[error==TRUE])/sum(n)
  )
  rates <- rates %>% spread(match,rate) %>% rename(
    actual = `FALSE`,
    beta = `TRUE`
  )
  left_join(rates, totals)
}

# Multiple plot function
#
# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
# - cols:   Number of columns in layout
# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
#
# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
# then plot 1 will go in the upper left, 2 will go in the upper right, and
# 3 will go all the way across the bottom.
#
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
  #  ref: http://www.cookbook-r.com/Graphs/Multiple_graphs_on_one_page_(ggplot2)/
```

```{r data}
if (!file.exists("../data/all-sigs.rds")) {
  files <- dir("../data/signatures", pattern="csv")
  all <- data.frame()
  for (file in files) {
    tmp <- read.csv(file=file.path(path="../data/signatures", file)) 
    if (length(grep("x",names(tmp)) > 0)) tmp <- tmp %>% select(-x)
    all <- rbind(all, tmp)
  }
  all <- all %>% filter(land1_id < land2_id)
  saveRDS(all, file="../data/all-sigs.rds")
} else {
  all <- readRDS("../data/all-sigs.rds")
}

all$coarse <- 1
errors <- rbind(errorrate(all, 0.001), 
                errorrate(all, 0.005), 
                errorrate(all, 0.01), 
                errorrate(all, 0.05))
```

```{r data-profiles}
if (!file.exists("../data/all-profiles.rds")) {
  files <- dir("../data/profiles/", pattern="csv")
  all <- data.frame()
  for (file in files) {
    tmp <- read.csv(file=file.path(path="../data/profiles", file)) 
    #print(names(tmp))
    if (length(grep("coarse",names(tmp))) == 0)
      tmp$coarse <- gsub(".*coar-(.*).csv", "\\1", file)
    if (length(grep("x",names(tmp)) > 0)) tmp <- tmp %>% select(-x)
    all <- rbind(all, tmp)
  }
  all <- all %>% filter(land1_id < land2_id)
  all <- all %>% mutate(coarse = as.numeric(gsub("pt", ".", coarse)))
  saveRDS(all, file="../data/all-profiles.rds")
} else {
  all <- readRDS("../data/all-profiles.rds")
}

errors2 <- rbind(errorrate(all, 0.001), 
                errorrate(all, 0.005), 
                errorrate(all, 0.01), 
                errorrate(all, 0.05))
```

# Introduction and Background

## Problem statement and motivation

@@

\citep{chumbley} found that the algorithm gives false-positives in slightly over 2 cases (or 2%) of the time and about 9 false-negatives (or 9%) for every 100 comparisons. \citep{hadler} on the other hand, in the improved version of the algorithm first described by \citep{chumbley} found false-positives for 0 cases and 3 false-negatives (or 6%) for 50 comparisons of knwon matches.
\begin{figure}
\centering
\includegraphics[width=\textwidth]{images/B6-B2-L6.png}
```{r, echo=FALSE, fig.width=8, fig.height=3, out.width='\\textwidth'}
land <- read.csv("../data/b6b2l6.csv")
land %>% ggplot(aes(x = y, y=resid)) +geom_line() +theme_bw()
```

@@

## The Chumbley Score Test

\hh{You need to distinguish between $x(t)$ and $x(t_1)$ - $x(t)$ is a spatial process for some $t$. $x(t_1)$ is the realization of the process in $t_1$. - Alternatively you could also follow the notation in Hadler et al on 'digitized toolmarks at a pixel level' - where our 'pixel' is the resolution of the confocal light microscope, i.e. a pixel corresponds to 0.645 microns.}

The Chumbley score algorithm takes input as two vectorized processes $x(t_1)$ and $y(t_2)$ which denote two sets of marks or striae. These marks or striae are potentially from two different bullets or two different toolmarks whose source needs to be identified as being same or different. The marks or striae are indexed by the pixel location where $t_1$ is for the first striae referred to as $x$ and $t_2$ is for the second striae which is referred as $y$. $t_1$ and $t_2$ need not be the same but are usually of similar lenghtsThe similarity is then judged by the algorithm on the basis of cross-correlation of a fixed and constant number consecutive pixels say $k$ taken from the two indexed marks $x(t_1)$ and $y(t_2)$ such that in theory $k$ remains smaller that length of the two striae or marks which is the same as $t_1$ and $t_2$. Depending on the what stage of the algorithm we are in, matching of different pixel lengths and locations is done, which at the end effectively compares all possible windows that would guarantee in quantifying the two marks or striae as coming from the same source or not, which also lets us assess the error rates by checking for a large number of cases.

The algorithm works in two phases, namely, an optimization step and a validation step, at the end of which a Mann Whitney U statistic is calculated. A pre-processing step to the algorithm is to choose a coarseness value which  is used as a parameter to the LOWESS smoothing function. The coarseness essentially gives the proportion of points which influence the smooth at each value, which means larger values lead to more smoothness. The LOWESS smoothing is applied to each of two sets of vectorized striae or marks $x(t_1)$ and $y(t_2)$, before proceeding to the algorithm.

@@

## Detailed algorithm

###Optimization step
The idea behind this step is to first identify the area of best agreement in the two toolmark data. Comparison window size is predefined by the user, which in case of screwdriver toolmarks was chosen by Chumbley et al and Hadler et al as around 10 percent of the length of the toolmarks, which was around 500. This window is henceforth referred to as Window of optimization.

@@

First, there are a very large number of cross-correlations calculated for the two series of striae or marks $x(t_1)$ and $y(t_2)$, for eg if the window of optimization was defined as 200 and toolmark pixel length is around 1000 then we have 1200-(200-1) = 1001. Can be thought of ordered arrangements without replacement where n = 1001 and r = 1.

@@

###Validation Step

####Same-shift: 
In this step a series of windows are chosen at random (originally by chumbley et al) and deterministically (by hadler et al), but at a common distance (rigid-shift) from the window identified as the region of best agreement in the vaildation step.
The correlation of these windows would be as intuitively assumed i.e. lower than the maximum correlation window (Optimization step), but the significance is that these same shift windows will still have large enough correlation values for two toolmarks or signatures that are in reality a match.

@@

```{r win-comparison, fig.width=8, fig.height=5, out.width='\\textwidth', fig.cap=" The two plots on the left show how the same shift behaves in case of a matching pair and the two plots on the right show how the different shift behaves in case of a matching pair."}
a<- matrix(NA, ncol = 3, nrow = 2)
toolmark.falsepositive<- data.frame(a)
colnames(toolmark.falsepositive)<- c("Classification", "Match", "Non-Match")
toolmark.falsepositive$Classification<- c("Match", "Non-Match")
toolmark.falsepositive$Match<- c(47,0)
toolmark.falsepositive$`Non-Match`<- c(3,50)
toolmark.falsepositive

asd<- all %>% filter(wo==120, wv==50, match == TRUE, land1_id == 1)

sigs_graphics<- readRDS("../data/sigs_generate_window.rds")

sigs_graphics.match<- as.data.frame(sigs_graphics[1])
sigs_graphics.nonmatch<- as.data.frame(sigs_graphics[2])
colnames(sigs_graphics.match)<- gsub("^.*\\.","", colnames(sigs_graphics.match))
colnames(sigs_graphics.nonmatch)<- gsub("^.*\\.","", colnames(sigs_graphics.nonmatch))

# Matching signature
#gi<-
#sigs_graphics.match<- sigs_graphics.match %>% filter(land_id == c(1,9))  
#sigs_graphics.match<- sigs_graphics.match %>% select(y, l30, land_id)
d1<- sigs_graphics.match[which(sigs_graphics.match$land_id==1,arr.ind = TRUE),]
d2<- sigs_graphics.match[which(sigs_graphics.match$land_id==19,arr.ind = TRUE),]
data1<-d1 %>%select(l30)
#data1<- data1$y
data2<- d2 %>%select(l30)
#data2<- data2$y
window_opt = 120 
window_val = 50
coarse = 1
data1<- matrix(unlist(data1))
data2<- matrix(unlist(data2))

  unity <- function(x) {x / sqrt(sum(x^2))} ## normalize columns of a matrix to make correlation computation faster

  ####################################################
  ##Clean the marks and compute the smooth residuals##
  ####################################################

  data1 <- matrix(data1[round((0.01*nrow(data1))):round(0.99*nrow(data1)),], ncol = 1)
  data2 <- matrix(data2[round((0.01*nrow(data2))):round(0.99*nrow(data2)),], ncol = 1)

  ##Normalize the tool marks
  y1 <- data1 - lowess(y = data1,  x = 1:nrow(data1), f= coarse)$y
  y2 <- data2 - lowess(y = data2,  x = 1:nrow(data2), f= coarse)$y

  ############################################
  ##Compute the observed maximum correlation##
  ############################################

  #####################
  ##Optimization step##
  #####################
  ##Each column in these matrices corresponds to a window in the respective tool mark
  y1_mat_opt <- matrix(NA, ncol = length(1:(length(y1) - (window_opt - 1))), nrow = window_opt)
  for(l in 1:(length(y1) - (window_opt - 1))){
    y1_mat_opt[,l] <- y1[l:(l+(window_opt - 1))]
  }
  y2_mat_opt <- matrix(NA, ncol = length(1:(length(y2) - (window_opt - 1))), nrow = window_opt)
  for(l in 1:(length(y2) - (window_opt - 1))){
    y2_mat_opt[,l] <- y2[l:(l+(window_opt - 1))]
  }

  ##Compute the correlation between all pairs of windows for the two marks
  ##Rows in the following matrix are mark 2, columns are mark 1
  y2_mat_opt <- apply(scale(y2_mat_opt), 2, unity)
  y1_mat_opt <- apply(scale(y1_mat_opt), 2, unity)
  corr_mat_opt <- t(y2_mat_opt) %*% y1_mat_opt ##correlation matrix
  max_corr_opt_loc <- which(corr_mat_opt == max(corr_mat_opt), arr.ind = TRUE) ##pair of windows maximizing the correlation

  s1.rect.opt <- data.frame(xmin=max_corr_opt_loc[1,2], xmax= max_corr_opt_loc[1,2]+window_opt, ymin=-Inf, ymax=Inf)
    s1.rect.sameshift1 <- data.frame(xmin=max_corr_opt_loc[1,2]-window_val, xmax= max_corr_opt_loc[1,2]- 2*window_val, ymin=-Inf, ymax=Inf)
       s1.rect.sameshift2 <- data.frame(xmin=max_corr_opt_loc[1,2]+window_opt+ window_val, xmax= max_corr_opt_loc[1,2]+ window_opt+2*window_val, ymin=-Inf, ymax=Inf)

  s2.rect.opt <- data.frame(xmin=max_corr_opt_loc[1,1], xmax= max_corr_opt_loc[1,1]-window_opt, ymin=-Inf, ymax=Inf)
    s2.rect.sameshift1 <- data.frame(xmin=max_corr_opt_loc[1,1]-window_opt-window_val, xmax= max_corr_opt_loc[1,1]-window_opt- 2*window_val, ymin=-Inf, ymax=Inf)
       s2.rect.sameshift2 <- data.frame(xmin=max_corr_opt_loc[1,1]+ window_val, xmax= max_corr_opt_loc[1,1]+ 2*window_val, ymin=-Inf, ymax=Inf)

p1<-  ggplot(data = d1, aes(x = y, y = l30)) + geom_line() + 
  geom_rect(data=s1.rect.opt, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "black",alpha=0.5, inherit.aes = FALSE)+
    geom_rect(data=s1.rect.sameshift1, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "red",alpha=0.1, inherit.aes = FALSE)+
    geom_rect(data=s1.rect.sameshift2, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "blue",alpha=0.1, inherit.aes = FALSE)+
  theme_bw() + labs(y = "Signature Land Id = 1")

 p2<-   ggplot(data = d2, aes(x = y, y = l30)) + geom_line() + 
   geom_rect(data=s2.rect.opt, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "black",alpha=0.5, inherit.aes = FALSE)+
    geom_rect(data=s2.rect.sameshift1, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "red",alpha=0.1, inherit.aes = FALSE)+
    geom_rect(data=s2.rect.sameshift2, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "blue",alpha=0.1, inherit.aes = FALSE)+
  theme_bw() + labs(y = "Signature Land Id = 9")
 ##########

 d1.rect.opt <- data.frame(xmin=max_corr_opt_loc[1,2], xmax= max_corr_opt_loc[1,2]+window_opt, ymin=-Inf, ymax=Inf)
    d1.rect.sameshift1 <- data.frame(xmin=max_corr_opt_loc[1,2]-2*window_opt, xmax= max_corr_opt_loc[1,2]-2*window_opt-window_val, ymin=-Inf, ymax=Inf)
       d1.rect.sameshift2 <- data.frame(xmin=max_corr_opt_loc[1,2]+window_opt+ window_val, xmax= max_corr_opt_loc[1,2]+ window_opt+2*window_val, ymin=-Inf, ymax=Inf)

  d2.rect.opt <- data.frame(xmin=max_corr_opt_loc[1,1], xmax= max_corr_opt_loc[1,1]-window_opt, ymin=-Inf, ymax=Inf)
    d2.rect.sameshift1 <- data.frame(xmin=max_corr_opt_loc[1,1]-window_opt-window_val, xmax= max_corr_opt_loc[1,1]-window_opt- 2*window_val, ymin=-Inf, ymax=Inf)
       d2.rect.sameshift2 <- data.frame(xmin=max_corr_opt_loc[1,1]+ 4*window_val, xmax= max_corr_opt_loc[1,1]+ 5*window_val, ymin=-Inf, ymax=Inf)

p3<-  ggplot(data = d1, aes(x = y, y = l30)) + geom_line() + 
  geom_rect(data=d1.rect.opt, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "black",alpha=0.5, inherit.aes = FALSE)+
    geom_rect(data=d1.rect.sameshift1, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "red",alpha=0.1, inherit.aes = FALSE)+
    geom_rect(data=d1.rect.sameshift2, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "blue",alpha=0.1, inherit.aes = FALSE)+
  theme_bw() + labs(y = "Signature Land Id = 1")

 p4<-   ggplot(data = d2, aes(x = y, y = l30)) + geom_line() + 
   geom_rect(data=d2.rect.opt, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "black",alpha=0.5, inherit.aes = FALSE)+
    geom_rect(data=d2.rect.sameshift1, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "red",alpha=0.1, inherit.aes = FALSE)+
    geom_rect(data=d2.rect.sameshift2, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "blue",alpha=0.1, inherit.aes = FALSE)+
  theme_bw() + labs(y = "Signature Land Id = 9")

multiplot(p1, p2, p3, p4, cols=2)
```

####Different Shift:
Primary reason for this substep is to give perspective to the correlation values of the same shift window correlation values.

@@

###U Statistic:

@@

Let ns and nd be the number of same shift and different shift windows
$N = n_{s} + n_{d}$

The mann whitney U statistic is given by
$U =\sum^{ns}_{i=1}R_{s}\left( i\right)$

@@

$\overline{U}= \dfrac{U-M}{\sqrt{V}}$

@@

$M = n_{s}\left(\dfrac{N+1}{2}\right)$

@@

$V = \dfrac{n_{s}n_{d}}{N\left(N-1\right)}\left[\Sigma^{n_{s}}R_{s}\left(i\right)^{2}+\Sigma^{n_{d}}R_{d}\left(j\right)^{2}\right] -\dfrac{n_{s} n_{d}\left(N+1\right)^{2}}{4\left(N-1\right)}$

## Potential limitations of the Chumbley Score

\hh{in the application to striation marks on bullet lands: smaller in width and curved - need to adjust parameter settings; something similar is done in \citet{afte-chumbley} for toolmark comparisons of slip-joint pliers. XXX work in }

@@

## Testing setup:
\hh{Using cross-validation setup to identify appropriate parameter settings for (a) signatures and (b) profiles directly}

Following on similar lines to the setup of toolmarks, the first step here is to first identify what difference does different window sizes of optimization and the validation step have, when adapting the toolmark method to bullets.

@@

Bullet signatures are chosen by
1. Filtering out Land_id for Profiles from the Hamby 44 and Hamby 252 data and removing 
all NA values
2. run_id = 3 is chosen, Seems like the signatures generated from this run_id give the closest
match. Different run_id's have some different settings for generating the signatures
(The level of smoothing does not seem to be one of them)

The bullet signatures when generated already included the LOWESS smoothing.
Therefore, the coarseness factor is set to 1 while running the chumbley_non_random()
which generates the same_shift, different_shift, U-Stat and P_value parameters.

window Optimization 320 Window Validation 50
```{r echo=FALSE}
all %>% filter(wo==320, wv==50) %>% xtabs(data=., ~signif+match)
```
window Optimization 120 Window Validation 50
```{r echo=FALSE}
all %>% filter(wo==120, wv==50) %>% xtabs(data=., ~signif+match)
```
window Optimization 80 Window Validation 50
```{r echo=FALSE}
all %>% filter(wo==80, wv==50) %>% xtabs(data=., ~signif+match)

```

# Results

## Signatures

@@

We used the adjusted Chumbley method as proposed in \citet{hadler} and implemented in the R package `toolmaRk` \citep{toolmark} on all pairwise land-to-land comparisons of the Hamby scans (a total of 85,491 comparisons) in the following manner:

```{r, echo=FALSE, results='asis'}
param_setting<- all %>% select(wo, wv) %>% group_by(wo, wv) %>% tally() %>% select(wo,wv)
#param_setting<- data.frame(matrix(param_setting, nrow = 16))
#t(param_setting)
#param_setting %>% tally() 
library(kableExtra)
knitr::kable(t(param_setting), digits=0, caption = "Overview of parameter settings used for optimization and validation windows for bullet land signatures.", format = "latex" ,booktabs=TRUE) %>% kableExtra::kable_styling(latex_options = c("scale_down"))
```

### Signatures

@@

```{r type2, fig.width=8, fig.height=5, out.width='\\textwidth', fig.cap="Type II error rates observed across a range of window sizes for optimization $wo$. For a window size of $wo = 120$ we see a drop in type II error rate across all type I rates considered. Smaller validation sizes $wv$ are typically associated with a smaller type II error."}
errors %>% filter(wv %in% c(30,50)) %>%
  ggplot(aes(x = wo, y = beta, colour=factor(alpha))) + 
  geom_point(aes(shape=factor(wv))) +
#  facet_grid(.~wo, labeller="label_both") + 
  geom_smooth(aes(group=alpha), se=FALSE, method="loess") +
  theme_bw() +
  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") +
  scale_shape_discrete("Size of validation\nwindow, wv") +
  xlab(expression("Window size for optimization, "~wo)) +
  ylab("Type II error rate")

labels = expression(P[M1](tilde(z)>0),P[M0](tilde(z)>0))
```

@@

Coming back to the database of Hamby-44 and Hamby-252 datasets, the run_id = 3 was used when applying the chumbley algorithm on the signatures. The run_id not only defines the level of smoothing but also signifies the chosen height at which the profiles were selected initially.
Another important aspect is the range of horizontal values (which is referred to as the y values in \citet{aoas}) in the signatures. Thes have already been pre-processed in the database to not include any grooves.

Therefore for the sake of comparison the run_id = 3 is still chosen so as to ensure that the horizontal values remain the same as that of the signatures. This also gives us profiles with the grooves removed.

\hh{XXX ideas should go to the previous section on setting up the study; in this section we focus on results only XXX}
The idea therefore is to first use these raw values of the profile directly in the chumbley algorithm, and see how the algorithm performs for different coarseness values (smoothing parameter as referred in the function LOWESS used in the chumbley algorithm)

>>>>

% !TeX program = pdfLaTeX
\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{graphicx,psfrag,epsf}
\usepackage{enumerate}
\usepackage{natbib}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage[table]{xcolor}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{hyperref}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}

\usepackage{url} % not crucial - just used below for the URL

%\pdfminorversion=4
% NOTE: To produce blinded version, replace "0" with "1" below.
\newcommand{\blind}{0}

% DON'T change margins - should be 1 inch all around.
\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-.5in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.3in}%
\addtolength{\topmargin}{-.8in}%

\newenvironment{definition}[1]% environment name 
{% begin code 
  \par\vspace{.75\baselineskip}\noindent 
  \textbf{Definition (#1)}\begin{itshape}% 
  \par\vspace{.5\baselineskip}\noindent\ignorespaces 
}% 
{% end code 
  \end{itshape}\ignorespacesafterend 
}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\begin{document}

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\if0\blind
{
  \title{\bf Adaption of the Chumbley Score to matching of bullet striation marks}

  \author{
        Ganesh Krishnan \thanks{The authors gratefully acknowledge \ldots{}} \\
    Department of Statistics, Iowa State University\\
     and \\     Heike Hofmann \\
    Department of Statistics and csafe, Iowa State University\\
      }
  \maketitle
} \fi

\if1\blind
{
  \bigskip
  \bigskip
  \bigskip
  \begin{center}
    {\LARGE\bf Adaption of the Chumbley Score to matching of bullet striation marks}
  \end{center}
  \medskip
} \fi

\bigskip
\begin{abstract}

\end{abstract}

\noindent%
{\it Keywords:} 3 to 6 keywords, that do not appear in the title
\vfill

\newpage
\spacingset{1.45} % DON'T change the spacing!

@@

\section{Introduction}\label{introduction}

\subsection{Problem statement and
motivation}\label{problem-statement-and-motivation}

@@

\citep{chumbley} found that the algorithm gives false-positives in
slightly over 2 cases (or 2\%) of the time and about 9 false-negatives
(or 9\%) for every 100 comparisons. \citep{hadler} on the other hand, in
the improved version of the algorithm first described by
\citep{chumbley} found false-positives for 0 cases and 3 false-negatives
(or 6\%) for 50 comparisons of knwon matches.

\begin{figure}
\centering
\includegraphics[width=\textwidth]{images/B6-B2-L6.png}

\begin{center}\includegraphics[width=\textwidth]{figures/unnamed-chunk-1-1} \end{center}

@@

\subsection{Existing solution for toolmarks: Chumbley
Score}\label{existing-solution-for-toolmarks-chumbley-score}

\hh{XXX anything to do with motivation has to come first - move this to the first paragraph of the intro XXX}
\hh{XXX data and results come after the intro to chumbley.}
\hh{XXX start with defining the data, i.e. we are working with random processes $x(t)$ and $y(t)$ ... you should also go into the coarseness parameter here - that is a pre-processing step before the optimization. }
\gk{done}

The Chumbley score algorithm takes input as two vectorized processes
\(x(t_1)\) and \(y(t_2)\) which denote two sets of marks or striae.
These marks or striae are potentially from two different bullets or two
different toolmarks whose source needs to be identified as being same or
different. The marks or striae are indexed by the pixel location where
\(t_1\) is for the first striae referred to as \(x\) and \(t_2\) is for
the second striae which is referred as \(y\). \(t_1\) and \(t_2\) need
not be the same but are usually of similar lenghtsThe similarity is then
judged by the algorithm on the basis of cross-correlation of a fixed and
constant number consecutive pixels say \(k\) taken from the two indexed
marks \(x(t_1)\) and \(y(t_2)\) such that in theory \(k\) remains
smaller that length of the two striae or marks which is the same as
\(t_1\) and \(t_2\). Depending on the what stage of the algorithm we are
in, matching of different pixel lengths and locations is done, which at
the end effectively compares all possible windows that would guarantee
in quantifying the two marks or striae as coming from the same source or
not, which also lets us assess the error rates by checking for a large
number of cases.

The algorithm works in two phases, namely, an optimization step and a
validation step, at the end of which a Mann Whitney U statistic is
calculated. A pre-processing step to the algorithm is to choose a
coarseness value which is used as a parameter to the LOWESS smoothing
function. The coarseness essentially gives the proportion of points
which influence the smooth at each value, which means larger values lead
to more smoothness. The LOWESS smoothing is applied to each of two sets
of vectorized striae or marks \(x(t_1)\) and \(y(t_2)\), before
proceeding to the algorithm.

@@

\subsection{Chumbley Algorithm and Chumbley
score}\label{chumbley-algorithm-and-chumbley-score}

\subsubsection{Optimization step}\label{optimization-step}

The idea behind this step is to first identify the area of best
agreement in the two toolmark data. Comparison window size is predefined
by the user, which in case of screwdriver toolmarks was chosen by
Chumbley et al and Hadler et al as around 10 percent of the length of
the toolmarks, which was around 500. This window is henceforth referred
to as Window of optimization.

@@

First, there are a very large number of cross-correlations calculated
for the two series of striae or marks \(x(t_1)\) and \(y(t_2)\), for eg
if the window of optimization was defined as 200 and toolmark pixel
length is around 1000 then we have 1200-(200-1) = 1001. Can be thought
of ordered arrangements without replacement where n = 1001 and r = 1.

@@

\subsubsection{Validation Step}\label{validation-step}

\paragraph{Same-shift:}\label{same-shift}

In this step a series of windows are chosen at random (originally by
chumbley et al) and deterministically (by hadler et al), but at a common
distance (rigid-shift) from the window identified as the region of best
agreement in the vaildation step. The correlation of these windows would
be as intuitively assumed i.e.~lower than the maximum correlation window
(Optimization step), but the significance is that these same shift
windows will still have large enough correlation values for two
toolmarks or signatures that are in reality a match.

@@

\begin{verbatim}
  Classification Match Non-Match
1          Match    47         3
2      Non-Match     0        50
\end{verbatim}

\begin{figure}

{\centering \includegraphics[width=\textwidth]{figures/win-comparison-1}

}

\caption{ The two plots on the left show how the same shift behaves in case of a matching pair and the two plots on the right show how the different shift behaves in case of a matching pair.}\label{fig:win-comparison}
\end{figure}

\paragraph{Different Shift:}\label{different-shift}

Primary reason for this substep is to give perspective to the
correlation values of the same shift window correlation values.

@@

\subsubsection{U Statistic:}\label{u-statistic}

@@

Let ns and nd be the number of same shift and different shift windows
\(N = n_{s} + n_{d}\)

The mann whitney U statistic is given by
\(U =\sum^{ns}_{i=1}R_{s}\left( i\right)\)

@@

\(\overline{U}= \dfrac{U-M}{\sqrt{V}}\)

@@

\(M = n_{s}\left(\dfrac{N+1}{2}\right)\)

@@

\(V = \dfrac{n_{s}n_{d}}{N\left(N-1\right)}\left[\Sigma^{n_{s}}R_{s}\left(i\right)^{2}+\Sigma^{n_{d}}R_{d}\left(j\right)^{2}\right] -\dfrac{n_{s} n_{d}\left(N+1\right)^{2}}{4\left(N-1\right)}\)

\subsection{\texorpdfstring{Potential limitations in the application to
striation marks on bullet lands: smaller in width and curved - need to
adjust parameter settings; something similar is done in
\citet{afte-chumbley} for toolmark comparisons of slip-joint
pliers.}{Potential limitations in the application to striation marks on bullet lands: smaller in width and curved - need to adjust parameter settings; something similar is done in  for toolmark comparisons of slip-joint pliers.}}\label{potential-limitations-in-the-application-to-striation-marks-on-bullet-lands-smaller-in-width-and-curved---need-to-adjust-parameter-settings-something-similar-is-done-in-for-toolmark-comparisons-of-slip-joint-pliers.}

@@

\subsection{Using cross-validation setup to identify appropriate
parameter settings for (a) signatures and (b) profiles
directly}\label{using-cross-validation-setup-to-identify-appropriate-parameter-settings-for-a-signatures-and-b-profiles-directly}

Following on similar lines to the setup of toolamrks, the first step
here is to first identify what difference does different window sizes of
optimization and the validation step have, when adapting the toolmark
method to bullets.

@@

Bullet signatures are chosen by 1. Filtering out Land\_id for Profiles
from the Hamby 44 and Hamby 252 data and removing all NA values 2.
run\_id = 3 is chosen, Seems like the signatures generated from this
run\_id give the closest match. Different run\_id's have some different
settings for generating the signatures (The level of smoothing does not
seem to be one of them)

The bullet signatures when generated already included the LOWESS
smoothing. Therefore, the coarseness factor is set to 1 while running
the chumbley\_non\_random() which generates the same\_shift,
different\_shift, U-Stat and P\_value parameters.

window Optimization 320 Window Validation 50

\begin{verbatim}
       match
signif  FALSE  TRUE
  FALSE 77564   412
  TRUE   4022   674
\end{verbatim}

window Optimization 120 Window Validation 50

\begin{verbatim}
       match
signif  FALSE  TRUE
  FALSE 79249   386
  TRUE   4523   854
\end{verbatim}

window Optimization 80 Window Validation 50

\begin{verbatim}
       match
signif  FALSE  TRUE
  FALSE 78817   450
  TRUE   4680   775
\end{verbatim}

\section{Results}\label{results}

\subsection{Signatures}\label{signatures}

@@

We used the adjusted Chumbley method as proposed in \citet{hadler} and
implemented in the R package \texttt{toolmaRk} \citep{toolmark} on all
pairwise land-to-land comparisons of the Hamby scans (a total of 85,491
comparisons) in the following manner:

\begin{table}

\caption{\label{tab:unnamed-chunk-5}Overview of parameter settings used for optimization and validation windows for bullet land signatures.}
\centering
\resizebox{\linewidth}{!}{\begin{tabular}[t]{lrrrrrrrrrrrrrrrrrrrrrrrrrrrrrrrrr}
\toprule
wo & 50 & 50 & 60 & 60 & 80 & 80 & 90 & 90 & 100 & 100 & 110 & 110 & 120 & 120 & 120 & 120 & 120 & 130 & 130 & 140 & 140 & 150 & 150 & 160 & 160 & 200 & 200 & 200 & 240 & 240 & 280 & 280 & 320\\
wv & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 10 & 20 & 30 & 50 & 60 & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 20 & 30 & 50 & 30 & 50 & 30 & 50 & 50\\
\bottomrule
\end{tabular}}
\end{table}

\subsubsection{Signatures}\label{signatures-1}

@@

\begin{figure}

{\centering \includegraphics[width=\textwidth]{figures/type2-1}

}

\caption{Type II error rates observed across a range of window sizes for optimization $wo$. For a window size of $wo = 120$ we see a drop in type II error rate across all type I rates considered. Smaller validation sizes $wv$ are typically associated with a smaller type II error.}\label{fig:type2}
\end{figure}

@@

Coming back to the database of Hamby-44 and Hamby-252 datasets, the
run\_id = 3 was used when applying the chumbley algorithm on the
signatures. The run\_id not only defines the level of smoothing but also
signifies the chosen height at which the profiles were selected
initially. Another important aspect is the range of horizontal values
(which is referred to as the y values in \citet{aoas}) in the
signatures. Thes have already been pre-processed in the database to not
include any grooves.

Therefore for the sake of comparison the run\_id = 3 is still chosen so
as to ensure that the horizontal values remain the same as that of the
signatures. This also gives us profiles with the grooves removed.

The idea therefore is to first use these raw values of the profile
directly in the chumbley algorithm, and see how the algorithm performs
for different coarseness values (smoothing parameter as referred in the
function LOWESS used in the chumbley algorithm)

\bibliographystyle{agsm}
\bibliography{bibliography}

\end{document}
